---
title: "week_04_assign_04"
author: "ashraf.said@etu.unige.ch"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
Cities = read.table("C:\\Users\\JEmre\\Downloads\\Cities.csv", sep = ",", header = TRUE)
head(Cities)
```


```{r}
str(Cities)
```


```{r}
class(Cities$City) 
class(Cities$Work)
class(Cities$Price)
class(Cities$Salary)
```
```{r}

# eliminate NA values
Cities.0nas = na.omit(Cities) # delete observations which contain at least one NA

#check the summaries again:
summary(Cities.0nas) # A basic description

mean(Cities.0nas$Work) 
median(Cities.0nas$Work)
quantile(Cities.0nas$Work, probs = c(0.25,0.75)) 
max(Cities.0nas$Work)
```



```{r}
par(mfrow=c(1,3))
boxplot(Cities$Work,col="lightsalmon1", main = "Work")
# with the default color changed to lightsalmon
boxplot(Cities$Price,col="mediumseagreen", main = "Price")
boxplot(Cities$Salary,col="goldenrod2", main = "Salary")
par(mfrow=c(1,1)) # back to the default setting
```

let us check densities plots 
```{r}
par(mfrow=c(1, 3)) 
plot(density(Cities.0nas$Work))
plot(density(Cities.0nas$Price))
plot(density(Cities.0nas$Salary))
```

Draw histograms of all the numerical (continuous) variables into a single graphical window. Use here as well the mfrow and col parameters to change the default settings
```{r}
par(mfrow=c(1,3)) # 3 figures arranged in a row
hist(Cities.0nas$Work, col = "lightsalmon1")
hist(Cities.0nas$Price, col = "mediumseagreen")
hist(Cities.0nas$Salary, col = "goldenrod2")
par(mfrow=c(1,1)) # back to the default setting
```


```{r}
library(vioplot)
par(mfrow=c(1 ,3)) # 3 figures arranged in a row
vioplot(Cities$Work, main = "Work", col = "lightblue")
vioplot(Cities$Price, main = "Price", col = "seagreen")
vioplot(Cities$Salary, main = "Salary", col = "pink")
par(mfrow=c(1 ,1)) # back to the default setting
```
Since the vioplot is the combination of boxplot and kernel density plot, we can draw the same conclusions as in 2 and 3. But we cannot identify the outlier from vioplot.


Next Compare, via QQ-plots, the empirical distribution of variables `Work, Price and Salary` separately with the Gaussian distribution and draw a reference line. Does the Gaussian distribution fit well?

```{r}
par(mfrow=c(1 ,3)) # 3 figures arranged in a row
qqnorm(Cities$Work, main="Work")
qqline(Cities$Work, col = "blue", lwd = 2)
qqnorm(Cities$Price, main="Price")
qqline(Cities$Price, col = "seagreen", lwd =2)
qqnorm(Cities$Salary, main="Salary")
qqline(Cities$Salary, col = "darkred", lwd = 2)
```
From the QQ-plots, we can conclude that the variable ”Work” and ”Price” are roughly normally distributed, but variable ”Salary” is not normally distributed since there is a clear departure in the tails.