---
title: "week_05_assign_05"
author: "ashraf.said@etu.unige.ch"
date: "`r Sys.Date()`"
output: html_document
---

let us first compute the probability corresponding to the below quantile using the *pnorm function* and the given normal distribution parameters 
```{r}
# P(X < 12)
quan_12 <- pnorm(q = 12, mean = 10, sd = sqrt(4))
quan_12
# P(X > 10)
quan_10 <- pnorm(q = 10 , mean =10 , sd = sqrt(4))
quan_10

#P( 10 < X < 12) = F(12) - F(10)
quan_12-quan_10
```
let us compute the 90th percentile of an expo distribution with rate parameter 1/2 , 1/4 and 2

```{r}
qexp(p = 0.9, rate = 1/2)
qexp(p = 0.9, rate = 1/4)
qexp(p = 0.9, rate = 2) 

```



```{r}
# P(X=2)
dpois(x =  2,lambda = 4)

# P(X <= 2)
# Careful with the definition of CDF: For <= 2, we use ppois() or we add P(X=0)+P(X=1)+P(X=2)
ppois(q = 2,lambda = 4, lower.tail = TRUE)

#dpois(x = 0, lambda = 4) + dpois(x = 1, lambda = 4) + dpois(x = 2, lambda = 4) alternatively 
sum(dpois(x = 0:2,lambda = 4))
```

let us generate some random numbers from normal with parameters 
1.5 and 2 as mean and standard deviation respectively 

```{r}
norm.sample <- rnorm(n = 200, mean = 1.5, sd = 2)
head(norm.sample)
max(norm.sample);min(norm.sample);range(norm.sample);norm.sample[76:79]
```


```{r}
plot(density(norm.sample), main = "Kernel Density")
```




```{r}
plot(density(norm.sample), main = "Kernel Density")
lines(norm.sample, dnorm(norm.sample, mean =1.5, sd = 2 ), col="red")

```


let us do some ordering or sorting the points 

```{r}
plot(density(norm.sample), main = "Kernel Density")
sorted.norm.sample = sort(norm.sample)
lines(sorted.norm.sample, dnorm(sorted.norm.sample,mean  = 1.5, sd = 2),
col="red")
```

note that above the black line represents the simulated data points and the red is the one which represents actually the theoretical normal distribution 
```{r}
mean(sorted.norm.sample);sd(sorted.norm.sample)
mean(norm.sample);sd(norm.sample)

```
As an example, we shall study the behavior of the sampling distribution of the median of a sample, median $\left(X_1, \ldots, X_n\right)$, when the observations are drawn from an exponential distribution, with the below CDF 


$$
F(x)=1-\exp (-\lambda x)
$$

To look at the sampling distribution of the median when $\lambda=1 / 2$ and sample size is 100, we simulate 500 samples: we proceed via standard for loop

```{r}
exp.median <- numeric(500)
for (i in 1:500){
  exp.median[i]<- median(rexp(100, rate = 1/2))
}
```

Explore graphically the sampling distribution of the above simulated median (with histograms, boxplots etc).

```{r}
par(mfrow=c(1,3))
hist(exp.median, probability = TRUE, col = "darkseagreen")
#lines(density(Exp.median),col="navyblue")
boxplot(exp.median, probability = TRUE, col = "salmon")
abline(h=1.39, col = "navyblue") #theoretical median we computed in point 1.
library(vioplot)
vioplot(exp.median, col = "lightblue")
```


let us check normality Q-Q plot

```{r}
par(mfrow=c(1,1))
qqnorm(exp.median)
qqline(exp.median)
# certainly looks like normal
```

Redoing the same steps as before for the 90-th quantile of the distribution rather than the 50-th percentile 

```{r}
exp_90q <- numeric(500)
for (i in 1:500){
  exp_90q [i]<- quantile(rexp(100, rate = 1/2),probs = 0.9)
} # where number of experiments = 500  and sample size per experiment is 100
```



```{r}
par(mfrow=c(1,3))
hist(exp_90q, probability = TRUE, col = "darkseagreen")

boxplot(exp_90q, probability = TRUE, col = "salmon")

abline(h=1.39, col = "navyblue") 

library(vioplot)
vioplot(exp_90q, col = "lightblue")
```

```{r}
par(mfrow=c(1,1))
qqnorm(exp_90q)
qqline(exp_90q)
```
